Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-06-07 jss/EpiCompare-paper/EpiCompare-paper.tex
major section 4 edits
also cleaned up end of section 3

\textbf{Comparisons and assessment} \textcolor{orange}{[[ NEW TEXT}

\textcolor{violet}{\sout{Finally,} Often}\footnote{\textcolor{violet}{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}}
the data analysis pipeline \textcolor{violet}{\sout{often}} ends with
plots, tables, and summary statistics that are used to
\sout{\textcolor{violet}{compare} \textcolor{orange}{and assess}}
\textcolor{violet}{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.}
In \pkg{EpiCompare} we provide a set of comparison and assessment tools
for model and simulation results that extend beyond the standard
performance metrics (e.g.~mean squared error or AIC) and into the lens
of time-invariant analysis. We have found \textcolor{violet}{that} these
tools \textcolor{violet}{\sout{to be} are} specifically applicable for
situations where only one season or cycle of an epidemic has occurred,
\textcolor{violet}{or is the object of interest}.

\footnote{\textcolor{violet}{[Ben says: This changed introduction was done to highlight that we can use prediction regions for mutiple types of comparisons.]}}\textcolor{violet}{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utilizing filamental representations of the model simulations and the true epidemic.}
{[}For three-state epidemic models, we provide the
\texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction \textcolor{violet}{region} to
the true epidemic trajectory
\textcolor{violet}{\sout{or comparing the prediction regions defined by two different models' simulations}}\footnote{\textcolor{violet}{[Ben says: this part will be highlighted in later sentences.]}}
can be done by eye. In \pkg{EpiCompare}, we also provide these
prediction regions for epidemic models with with more than three states.
The functions \texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}){]}.
\textcolor{violet}{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the}
\texttt{hausdorff\_dist()}
\textcolor{violet}{function to calculate the Hausdorff distance between multiple prediction regions, when visual comparison is not possible.}

{[}We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}.{]}
\textcolor{violet}{We provide implementations of a few distance-based score functions that capture how "reasonable" an epidemic is relative to other epidemics, and these scores can be turned into an extremeness measure with}
\texttt{mean(sim\_scores\ \textgreater{}\ truth\_score)}.
{[}Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.{]}

@@

{[}{[}\textcolor{red}{ OLD TEXT}

One tool we provide to assess models is through the creation of
geometric prediction regions, which are useful when we treat epidemics
like filaments. If there is a set of simulated epidemics from a model,
we can create a geometric prediction region for the expected trajectory
of the epidemic in the state space. For three-state epidemic models, we
provide the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done by eye. In \pkg{EpiCompare}, we also
provide these prediction regions for epidemic models with with more than
three states. The functions \texttt{create\_convex\_hull\_structure()}
and \texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}) and the ability to
assess the Hausdorff distance between prediction regions based on
simulations from different model (\texttt{hausdorff\_dist()}). The
vignettes for \texttt{EpiCompare} go into much more mathematical detail
about how these functions can be used in conjunction with time-invariant
analysis.

We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}. Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.

{[}{[}\textcolor{red}{Really OLD TEXT}

Finally, in \pkg{EpiCompare} we provide a set of comparison and
assessment tools for model and simulation results that extend beyond the
standard performance metrics (e.g.~mean squared error or AIC).
Corresponding\footnote{\textcolor{violet}{[Ben says: "corresponding to" doesn't seem to be the correct phrase here. If we wanted to use "corresponding" I might imagine saying "corresponding to the araguments for time-invariant analysis in Section 2.2, EpiCompare provides such tools." I'm open to keeping it if you think it's the perfect word choice.]}\textcolor{orange}{"To support the ideas presented in Sec."?}}
to Section \ref{sec:beyond-r0-sir}, \pkg{EpiCompare} provides a set of
time-invariant tools to compare and evaluate epidemic models and
simulations. We have found these tools to be specifically applicable for
situations where only one ``season'' or
``cycle''\footnote{\textcolor{violet}{[Ben says: I think we can remove the paratheses here.]} \textcolor{orange}{accepted}}
of the epidemic has occurred.

\footnote{\textcolor{violet}{[Ben says: multiple parts of this paragraph might benefit with capturing a bit more of the motivation in the rewrite of Section 2 for the tools we're presenting.]}\textcolor{orange}{motivate better}}One
tool we provide to assess models is through the creation of geometric
prediction regions, which are useful when we treat epidemics like
filaments. If there is a set of simulated epidemics from a model, we can
create a geometric prediction region for the expected trajectory of the
epidemic in the state space. For three-state epidemic models, we provide
the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done
\textcolor{violet}{by eye.\sout{with the 'eye test.'}\textcolor{orange}{I'm not saying 'no' but I'd do it in an appendix and not the main text.  I don't think equations belong in Section 3.}}
In \pkg{EpiCompare} we also provide these prediction regions for
epidemic models with with more than three states. The functions
\texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}) and the ability to
assess the Hausdorff distance between prediction regions based on
simulations from different model (\texttt{hausdorff\_dist()}).

\footnote{\textcolor{violet}{[Ben says: random thought (not needed) - maybe we do provide some equations for these measures and also maybe mathematically express how we'd examine extremeness?]}}We
also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}. Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations. {]}{]}

@@

\footnote{\textcolor{violet}{[Ben says: Generally I found this whole section to be pretty passive and not well motivated on why we'd actually do the analysis.]}}
To conclude our paper, we demonstrate the capabilities of
\pkg{EpiCompare} with a complete data analysis of a measles outbreak in
1861-1862 Germany. Specifically, we demonstrate how tools in
\pkg{EpiCompare} can be used in each step of the data analysis pipeline
\footnote{\textcolor{violet}{[Ben says: based on an earlier read of this section, I might suggest something like "and streamline the analysis process" - is that a selling point you want to highlight?]}}
\textcolor{violet}{\sout{(see Figure \ref{fig:pipeline})}}\footnote{\textcolor{violet}{[Ben says: It's unclear why you want to highlight the figure again - could you be clearere on that? / what does it add to the conv/ why should they reference it?]}}.
Additionally, we highlight how time-invariant analysis (see Section
\ref{sec:time-invariant}) can be used to enhance understanding of an
outbreak\footnote{\textcolor{violet}{[Ben says: I'm not sure this is necessary to state. Additionally - is it really true do we highlight this?]}}.

@@

{\centering \includegraphics{Figs/unnamed-chunk-7-1}

@@

\caption{\label{fig:cifs}Empirical cumulative incidence functions of prodrome (symptom) onset and measles rash appearance.  We see that there is approximately a a constant lag between the two curves.}\label{fig:unnamed-chunk-7}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics{Figs/unnamed-chunk-9-1}

@@

\caption{\label{fig:hag-tern-raw}Time invariant view of the Hagelloch epidemic where we view the individuals in Susceptible, Infectious, or Recovered states.  We see there are two peaks of infection (the vertical axis).}\label{fig:unnamed-chunk-9}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics{Figs/unnamed-chunk-10-1}

@@

\caption{\label{fig:tern-class-data}Time invariant outbreak curves for the three class groups.  The pre-school class has a distinct peak of infection whereas the peak infection point for the other two classes are less well defined.}\label{fig:unnamed-chunk-10}
\end{figure}
\end{CodeChunk}

@@

\hypertarget{modeling-and-simulation}{%
\subsection{Modeling and simulation}\label{modeling-and-simulation}}

@@

The next step is to compare the simple SIR model to the EpiModel SIR
model. We
provide\footnote{\textcolor{violet}{[Ben says: the phrase "We provide" is very passive / distance from the current demonstration at hand. Moreover section 3 already phrases things this way.]}}
the function \code{fortify_aggregate()}, which can take objects from
specialized classes of modeling output
\textcolor{orange}{(like those made by \code{netsim()})} and transform
it into a tidy-style data frame.

@@

{\centering \includegraphics{Figs/unnamed-chunk-17-1}

@@

\caption{\label{fig:hag-simple-sir}  Original Hagelloch SIR data (black) along with 90\% prediction band and actual simulation paths from the Simple SIR and the EpiModel SIR models.}\label{fig:unnamed-chunk-17}
\end{figure}
\end{CodeChunk}

>>>>

\textbf{Comparisons and assessment}

The last step of the data analysis pipeline often ends with plots,
tables, and summary statistics that are used to assess model performance
and compare across models or simulations. In \pkg{EpiCompare} we provide
a set of comparison and assessment tools for model and simulation
results that extend beyond the standard performance metrics (e.g.~mean
squared error or AIC) and into the lens of time-invariant analysis. We
have found that these tools are specifically applicable for situations
where only one season or cycle of an epidemic has occurred or is the
object of interest.

The first set of of tools surround the creation of prediction regions.
We can create a prediction regions from model simulations to examine if
our model simulations capture the true epidemic trajectory. We do so in
a time-invariant way and utilizing filamental representations of the
model simulations and the true epidemic. For three-state epidemic
models, we provide the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction region to the true epidemic
trajectory can be done by eye. In \pkg{EpiCompare}, we also provide
these prediction regions for epidemic models with with more than three
states. The functions \texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}){]}. We can also use
these prediction regions to visually or mathematically compare how
similar two sets of simulations are. In \pkg{EpiCompare} we provide the
\texttt{hausdorff\_dist()} function to calculate the Hausdorff distance
between multiple prediction regions, when visual comparison is not
possible.

We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}.{]} We provide implementations of a few
distance-based score functions that capture how ``reasonable'' an
epidemic is relative to other epidemics, and these scores can be turned
into an extremeness measure with
\texttt{mean(sim\_scores\ \textgreater{}\ truth\_score)}.
{[}Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.

\textcolor{orange}{I have substantially overhauled section 4.  The older versions are kept after the entirety of this section but note that much of the text (but not the structure) of this section is different.  I've tried to better motivate why we use the epicompare functions to answer our questions of interest.}

\textcolor{orange}{[[NEWEST TEXT}

To conclude our paper, we demonstrate how \pkg{EpiCompare} can
streamline the data analysis process with a case study of a measles
outbreak in 1861-1862 Germany. With the help of \pkg{EpiCompare}, we can
answer questions such as is a SIR model a good fit for the data, does
the outbreak spread differently within the different school classes of
the children, and can incorporating an underlying network structure
enhance model fit. We first begin with some context for the epidemic.

\hypertarget{background-for-1861-1862-measles-epidemic}{%
\subsection{Background for 1861-1862 measles
epidemic}\label{background-for-1861-1862-measles-epidemic}}

We analyze an outbreak of measles in the town of Hagelloch, Germany from
1861-1862, a data set organized by \cite{pfeilsticker1863}. The data was
later made visible by \cite{oesterle1992} and made available in an
\proglang{R} by \cite{surveillance2017}. In this outbreak, 188 children
were infected with measles over the course of three months. This data
set includes a rich collection of features including day of measles
rash, age, sex, school class, household and household location, and
alleged infector of each child. We show a subset of the data in Table
\ref{tab:hags-people}. We are particularly interested if and how group
structure plays a role in the spread of measles.

\hypertarget{pre-processing-and-eda}{%
\subsection{Pre-processing and EDA}\label{pre-processing-and-eda}}

We begin our analysis by first examining and transforming the raw data
(\code{hagelloch_raw}), which is individual-level data. With
\pkg{EpiCompare}, we can quickly determine how the epidemic progresses
through the population at an aggregate level. By specifying the time of
infection (\code{tI}) and the time of recovery (\code{tR}), the function
\code{agents_to_aggregate()} calculates the number susceptible,
infectious, and recovered individuals at each time step. Once
aggregated, we can plot the SIR values through a time-invariant lens
using \pkg{ggplot2} and \pkg{ggtern} functions (as shown in Fig.
\ref{fig:hag-tern-raw}) or with our custom \code{geom},
\code{geom_aggregate()}, which takes the raw agent data as input.

{\centering \includegraphics{Figs/unnamed-chunk-4-1}

\caption{\label{fig:hag-tern-raw}Time invariant view of the Hagelloch epidemic where we view the individuals in Susceptible, Infectious, or Recovered states.  We see there are two peaks of infection (the vertical axis).}\label{fig:unnamed-chunk-4}
\end{figure}
\end{CodeChunk}

In Figure \ref{fig:hag-tern-raw}, we can focus on the infections over
time by analyzing the vertical axis. Specifically, we see two peaks of
infection. This is interesting because the SIR model, which is sometimes
used to model the spread of measles, generally has one defined peak of
infection. We may wonder if the two peaks in the observed data may be
due to random noise or if a model more complex than the simple SIR is
needed to adequately capture these two peaks.

Previous study tells us that measles outbreaks are often associated with
children within the same grade level, and we examine if this is the case
here. By combining the function \code{facet_wrap()} with
\code{geom_aggregate()} we can easily analyze this scenario,

{\centering \includegraphics{Figs/unnamed-chunk-5-1}

\caption{\label{fig:tern-class-data}Time invariant outbreak curves for the three class groups.  The pre-school class has a distinct peak of infection whereas the peak infection point for the other two classes are less well defined.}\label{fig:unnamed-chunk-5}
\end{figure}
\end{CodeChunk}

Immediately in Fig. \ref{fig:tern-class-data}, we see that time
invariant infection curve is different for the pre-school class compared
to the 1st class. In the 1st class, we see about 95\% of the class
become infected and less than 10\% of them having recovered, which may
be indicative of a super-spreading event. This suspicion is further
supported in that 26 of the 30 1st class students have been reportedly
infected by the same individual. We now have some evidence that class
structure may play a role in the spread of infection. We can further
analyze this claim with modeling and simulation.

\hypertarget{modeling-and-simulation}{%
\subsection{Modeling and Simulation}\label{modeling-and-simulation}}

We first try to model the Hagelloch data with a baseline stochastic SIR
model, which we refer to as the `simple SIR.' In our full vignette
\textcolor{orange}{(available online),} we show how to fit this simple
SIR model via maximum likelihood, a common approach used to fit
parameters, and simulate from the model with those best fit parameters.
Our function \code{simulate_agents()} (or \code{simulate_SIR_agents()})
generates individual level data according to discrete-time multinomial
draws, which depend on the number of individuals in each state at the
previous time step and a matrix of transition probabilities. For
example, the below code generates 100 simulations of an outbreak of a
disease with one initial infector in a population of \(n= 188\)
individuals, a scenario analogous to the actual outbreak.

The result of our simulation is the object \code{agents} which is a
18800 \(\times\) 5 data frame, which details the time of entry into the
\(S\), \(I\), and \(R\) states for a given simulation. Ultimately, we
would like to know if this simple SIR model is a good fit to the data,
especially in comparison to a SIR model where we can incorporate the
class structure.

To fit a more complex SIR model with a network structure, we use package
\pkg{EpiModel} \citep{Jenness2018}. The below code sets up a network of
individuals (which includes class as a variable) and then simulates
infections over this network.

The output of this network model is \code{epimodel_sir}, an object of
class \code{netsim}, which contains a plethora of modeling information.
In the following section, we will use the capabilities of
\pkg{EpiCompare} to streamline the process of comparing the two models
contained in the objects \code{agg_model} (the simple SIR) and
\code{epimodel_sir} (the network SIR model).

The next step is to compare the simple SIR model to the EpiModel SIR
model. The \pkg{EpiCompare} function \code{fortify_aggregate()}, takes
in an object from specialized classes of modeling output
\textcolor{orange}{(like those made by \code{netsim()})} and transforms
it into a tidy-style data frame.

With the two modeling objects both in the same format, we can then
compare the models side-by-side. The results are shown in Figure
\ref{fig:hag-simple-sir}, where a 90\% prediction region is estimated
for the two models. For the Simple SIR model, we see that while the
prediciton region covers the data fairly well, the prediction region
clearly misses the second peak of infection. This indicates that the
simple SIR is not a good fit to our data. We also see that the
prediction region is very large, covering up a large area of the ternary
plot. Together, this indicates that the simple SIR model produces a
biased model with a large amount of variance. On the other hand, for the
EpiModel network model, we see that the prediction region covers the
data quite well and takes up less area. With \pkg{EpiCompare}, we can
see that the model using the class structure is a better fit to the
outbreak than the simple SIR model.

{\centering \includegraphics{Figs/unnamed-chunk-12-1}

\caption{\label{fig:hag-simple-sir}  Original Hagelloch SIR data (black) along with 90\% prediction band and actual simulation paths from the Simple SIR and the EpiModel SIR models.}\label{fig:unnamed-chunk-12}
\end{figure}
\end{CodeChunk}

The prediction region generated by the network model covers the observed
data well, but that does not mean that individual filaments generated
from the network model are good fits to the observed data. We can
further examine the model fits by incorporating these filaments into our
visualization. In Fig. \ref{fig:hag-simple-sir} we show the indivdual
filaments generated from the two sets of models as gray lines. Examining
these, we see that the individual lines typically only have one defined
peak, whereas the data certainly looks like it has two distinct peaks,
likely caused by our assumed super-spreader event.

We can also examine these filaments quantitatively by using functions
that take the distance between the filaments with the observed data. In
the below code, we transform the simulations to a more
computationally-friendly format with the function
\code{filament_compression()}. Following that, we calculate the distance
between the simulated filaments and the observed filament with the
function \code{dist_matrix_innersq_direction()} and calculate the
probability of the truth with respect to those simulations with the
function \code{compare_new_to_rest_via_distance()}. The estimated
pseudo-density of the observed epidemic (relative to the simulations
from either model) is much less likely then \textbf{any} of the
simulations (reported in Table \ref{tab:hags-extreme}). This indicates
that neither of two SIR models are good fits to the data at the filament
level.

\begin{CodeChunk}
\begin{CodeInput}
R> simple_sir <- both_models %>% filter(Type == "Simple SIR") %>%
+   rename(S = "X0", I = "X1", R = "X2") %>%
+   select(Type, sim, t, S, I, R)
R> 
R> hagelloch_sir2 <- hagelloch_sir %>%
+   rename(t = "time") %>%
+   mutate(Type = "true observation",
+          sim = 0) %>%
+   select(Type, sim, t, S, I, R)
\end{CodeInput}
\end{CodeChunk}

\caption{\label{tab:hags-extreme}The extremeness of the true simulations based on comparing pseudo-density estimates between true vs simulated curves}
\centering
\begin{tabular}[t]{l>{\raggedleft\arraybackslash}p{6cm}>{\raggedleft\arraybackslash}p{6cm}}
\toprule
Type & simulations-based estimated pseudo-density & proportion of simulations with lower estimated pseudo-density\\
\midrule
Simple SIR & 0.0036733 & 0.00\\
EpiModel SIR & 0.0118283 & 0.03\\
\bottomrule
\end{tabular}
\end{table}

In conclusion, \pkg{EpiCompare} allows us to fully examine this outbreak
at every step in the data analysis pipeline (see Fig.
\ref{fig:pipeline2}) in a streamlined fashion. With EDA, we saw evidence
that class structure may be important in the spread of of measles. We
then compared a baseline simple SIR model to a more complicated SIR
model which incorporated a network structure which included the class
structure. Based on the prediction regions generated from these models,
we saw that the network model fit the data much better than the simple
SIR model. However, when we examined the individual filaments generated
by the network model, we found that the data is highly unlikely to be
generated from such a model. For further analysis, we would recommend
looking into models that can more accurately capture super-spreading
events based on the observation that one child was allegedly responsible
for nearly all of his classmates' infections. Overall, this analysis
demonstrates how \pkg{EpiCompare} aids in the data analysis pipeline for
both novice and expert practitioners and coders alike.

@@

@@

\footnote{\textcolor{violet}{[Ben says: Generally I found this whole section to be pretty passive and not well motivated on why we'd actually do the analysis.]}}
To conclude our paper, we demonstrate the capabilities of
\pkg{EpiCompare} with a complete data analysis of a measles outbreak in
1861-1862 Germany. Specifically, we demonstrate how tools in
\pkg{EpiCompare} can be used in each step of the data analysis pipeline
\footnote{\textcolor{violet}{[Ben says: based on an earlier read of this section, I might suggest something like "and streamline the analysis process" - is that a selling point you want to highlight?]} \textcolor{orange}{check out new re-write.  new goal is to motivate question of hagelloch and then show how epicompare helps}}
\textcolor{violet}{\sout{(see Figure \ref{fig:pipeline})}}\footnote{\textcolor{violet}{[Ben says: It's unclear why you want to highlight the figure again - could you be clearere on that? / what does it add to the conv/ why should they reference it?]}}.
Additionally, we highlight how time-invariant analysis (see Section
\ref{sec:time-invariant}) can be used to enhance understanding of an
outbreak\footnote{\textcolor{violet}{[Ben says: I'm not sure this is necessary to state. Additionally - is it really true do we highlight this?]}}.

@@

{\centering \includegraphics{Figs/unnamed-chunk-22-1}

@@

\caption{\label{fig:cifs}Empirical cumulative incidence functions of prodrome (symptom) onset and measles rash appearance.  We see that there is approximately a a constant lag between the two curves.}\label{fig:unnamed-chunk-22}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics{Figs/unnamed-chunk-24-1}

@@

\caption{\label{fig:hag-tern-raw}Time invariant view of the Hagelloch epidemic where we view the individuals in Susceptible, Infectious, or Recovered states.  We see there are two peaks of infection (the vertical axis).}\label{fig:unnamed-chunk-24}
\end{figure}
\end{CodeChunk}

@@

{\centering \includegraphics{Figs/unnamed-chunk-25-1}

@@

\caption{\label{fig:tern-class-data}Time invariant outbreak curves for the three class groups.  The pre-school class has a distinct peak of infection whereas the peak infection point for the other two classes are less well defined.}\label{fig:unnamed-chunk-25}
\end{figure}
\end{CodeChunk}

@@

\hypertarget{modeling-and-simulation-1}{%
\subsection{Modeling and simulation}\label{modeling-and-simulation-1}}

@@

\hypertarget{post-processing-and-comparison-1}{%
\subsection{Post-processing and
comparison}\label{post-processing-and-comparison-1}}

The next step is to compare the simple SIR model to the EpiModel SIR
model. We
provide\footnote{\textcolor{violet}{[Ben says: the phrase "We provide" is very passive / distance from the current demonstration at hand. Moreover section 3 already phrases things this way.]}\textcolor{orange}{more of sentence 2 of ben than 1.}}
the function \code{fortify_aggregate()}, which can take objects from
specialized classes of modeling output
\textcolor{orange}{(like those made by \code{netsim()})} and transform
it into a tidy-style data frame.

@@

{\centering \includegraphics{Figs/unnamed-chunk-32-1}

@@

\caption{\label{fig:hag-simple-sir}  Original Hagelloch SIR data (black) along with 90\% prediction band and actual simulation paths from the Simple SIR and the EpiModel SIR models.}\label{fig:unnamed-chunk-32}
\end{figure}
\end{CodeChunk}
